Ontology Matching with word2vec

### 2016-06-22 Paper/Draft.tex
Shuffled sections around

 \section*{Summary}
 %TODO somewhere else: explain why one large ontology won't work
 %TODO Definitions
 %TODO Word index
 \tableofcontents
 \newpage

 \pagenumbering{arabic}
 \section{Introduction} \label{intro}
 It is a Friday afternoon and you realise you want to watch a movie with some friends this weekend. So, what do you do? To achieve this goal you must take four steps.

@@

 The most amazing thing about this story, is that it is already possible with the current technology. The algorithms to convert for example cinema website text to a computer-readable format exist. Many open databases already contain a lot of structured data that can be used to reason about data. And algorithms to combine these piece of information to make decisions already exist and work.

@@

 \subsection{Short Context}
 In this section, I will give a short summary of the context for the problem. This will enable the reader to put the problem in perspective as well as understand the reasoning behind the solution to the problem. Everything that is mentioned in this section will be explained more in-depth further on.

@@

 \subsection{Overview}
 This thesis is structured as follows: firstly, I will describe the context of the problem, which includes the Semantic Web, Ontologies, Ontology matching and other relevant research. Secondly, I will outline the problem and research questions. Thirdly, I will describe the method I used to develop the solution, which includes the idea, challenges and considered implementations. The fourth section after this describes methods I used to evaluate the solution, with the fifth section showing the results, the sixth section the conclusions and the last and seventh section the discussion.

\newpage
\section{Context}
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on Ontology matching, will be discussed. Secondly, the concept of the Ontology itself will be addressed. Thirdly, current alignment strategies and their their problems are discussed. Lastly, other related research and technologies will be discussed. Later sections will refer back to the concepts discussed in this section.
 \subsection{The Semantic Web}
 The first concept is the Semantic Web, also called Web 3.0, a vision of a Web that enhances the User Experience in ways that the current Web can not provide. Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and limited interconnectivity.\footnote{For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest.} It is possible to link to content from other hubs using URLs, but this moves the user to another hub, rather than connecting the content present in the hubs.
The Semantic Web, on the other hand, contains sources of \textit{information} that can be combined whenever the user needs it. To stick with the social media example, a person can create a piece of content, say a picture of a tree, and share it with his friends. One of his friends then can share this piece of content with a group of people that likes nature pictures. However, everyone can access all references someone makes. So if a person from the nature group notices the tree is ill and comments on this, the original poster can see this comment and act accordingly. On the current web the friend who shared the picture with the group has to be contacted directly by the commenter and then has to relay the message manually.

@@

 \subsubsection{Uses of Ontologies}
 %Beschrijf bottom-up ontologie maken.
 There are three main reasons Ontologies are used. Firstly, it allows querying languages to optimize the search process. Secondly, it allows expansion of information on a node. Thirdly, it allows for consistency checking a Knowledge Base. Lastly, it allows merging of knowledge bases.

@@

 \subsection{Current alignment strategies} \label{strategies}
 Many different alignment strategies have already been developed. All strategies follow the same two-step approach. The two steps are independent and as such different methods can be used interchangeably. The first step is to generate an initial correspondence set, where correspondences between nodes are found based on just their labels and meta-data. The second step is to use this initial set and the structure of the ontology to find more correspondences. These steps are called the terminological and the structural steps. Some strategies also use extra steps like the extensional and semantic steps, which use vector space models and inference, respectively\cite{matchingchallenges}.

@@

 Factorie is written in scala, a relatively new language that interfaces with java. See \ref{scala}. This allows for interoperability between factorie and AgreementMaker. Any algorithm that is implemented in factorie or on top of factorie can therefore be used as a matching algorithm in AgreementMaker.

@@

 \newpage
 \section{Problem}

 In summary, ontology aligning is required for multi-domain communication but poses a number of problems that make it difficult for concepts, the building blocks of ontologies, to be matched.
 \subsection{Research questions}
 The main question that I will answer with this work is the following research question:

 \subsection{Hypotheses} 
 \begin{center}
 \textit{This specific matcher that uses a distributed representation of concepts can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center}

 The first three points will be shown in \ref{proofs}, whereas the last point is shown in \ref{conclusions}.

 \begin{center} 
 \textit{The distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing terminological and structural matchers.}
 \end{center}
 \begin{center}
 \textit{The distributed representations matcher can perform well in a domain it was not trained for.}
 \end{center}

 These hypotheses are true for distributed representations matcher if they are true for one distributed representation matcher. Therefore to test the hypotheses, I must test them for every distinct implementation of the matcher.

 \subsection{Justification}
 On the Semantic Web, many relatively small knowledge bases will be combined temporarily so that it is possible to reason over the combined knowledge of those knowledge bases. To combine these knowledge bases, their ontologies first need to be aligned. This is an act that will happen very often on the Semantic Web, since every interaction with the Semantic Web will usually require information from multiple domains. Large domains may be stored over multiple devices, which can be viewed as different knowledge bases which also need to be aligned. Therefore, it is save to say that efficient ontology matching will be very valuable.

 Since most of the Semantic Web will consist of Aristotelian ontologies, some of the information it contains can be unreliable and even wrong. Therefore, ontology matching that is capable of handling this uncertain data is required.

 However, ontology matching is also used outside of the semantic web. It is already in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful in these situations, not just on the Semantic Web.

 \newpage
 \section{Method}

@@

 \newpage
 \section{Evaluation}
 %https://hal.inria.fr/hal-00917910/document
 The goal of the research is to find an ontology alignment algorithm $A$ and prove that it performs better than current techniques. However, as current techniques may combine many different algorithms and sources of information, maybe such a comparison is not fair, nor is a comparison with individual algorithms as they may be optimised for being combined. So, instead we may want to see if performance is improved when adding $A$ to an ensemble of algorithms $E=(A_1,..,A_n)$.
 \subsection{Evaluation methods} \label{Evaluation methods}
 The evaluation should be done using a performance measure $P(Al,T)$ that takes the alignment an algorithm produces given an ontology and the ground truth. If both the alignment and truth are considered sets of alignments, i.e. $Al=\{a_1,..,a_n\}$ and $T=\{a'_1,..,a'_m\}$, then the True Positives $TP=\{a|a\in Al, a\in T\}$, False Positives $FP=\{a|a\in Al, a\notin T\}$ and False Negatives $FN=\{a|a\in T, a\notin Al\}$ determine the performance of the alignment. The True Negatives do not matter, as there are extremely many, and the goal of ontology matching is to find True Positives. If we only care about True Positives, the Jaccard similarity is a prime candidate for the performance measure.

@@

 It is in principle possible to weigh False Positives and False Negatives differently, but this is not done in practice. To make the results of this study easily comparable to others, equal weights will be used.
 \subsection{Data}
 The data used comes from a list of data sets that the OAEI provides. % Zoek paper, misschien http://www.sciencedirect.com/science/article/pii/S1570826813000188
 %Hoeveel preprocessing moest ik doen
 Every data set contains two ontologies that need to be aligned and a ground truth alignment. The goal is straightforward: find the alignment given only the ontologies and whichever outside resources are needed.

@@

 \subsection{Precision versus recall}
  \subsubsection{Mice and men data}
  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/untrained_ROC}
   \caption{The precision and recall trade-off for the results of the untrained Node2vec. $AP = 0.04194$}
   \label{ROC_trained}
  \end{figure}

  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/Generic_ROC}
   \caption{The precision and recall trade-off for the results of the Word2vec algorithm. $AP: 0.04636$}
   \label{ROC_trained}
  \end{figure}

@@

  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/Alternative_ROC}
   \caption{The precision and recall trade-off for the results of the alternative DeepWalk algorithm. $AP = 0.04193$}
   \label{ROC_trained}
  \end{figure}
   \subsubsection{Medical data}
   TBD

 \newpage
 \section{Interpretation}
  The system underperforms on the provided data. This can have one or more of the following four causes:
  \begin{itemize}
  \item The data breaks the considered algorithms.
  \item There are bugs in the developed system unrelated to the algorithms.
  \item The considered context extraction does not work.
  \item The considered algorithms are not suited for Ontology matching.
  \end{itemize}

  \subsection{The data}
  If the data set is the cause of the underperformance, logically switching to another data set will improve performance over the baseline. Possible reasons the data set causes underperformance include a too small data set or a data set having a structure that does not work with the algorithm.

  The algorithm has been run on multiple data sets with no significant difference in performance, therefore the data is not the cause of the underperformance.
  \subsection{The system}
  If any of the steps in the software system contains an error, the whole system may fail. This could cause all algorithms to learn badly or the results to be misinterpreted. The system consists of a data loader, preprocessor, context extractor, the algorithm, the matcher and multiple visualizers.

@@

  As such, it is safe to conclude that the system has been thoroughly tested and is not faulty.
  \subsection{Context extraction}
  The context extraction works, as described in Section \ref{sentences}.
  \subsection{The algorithm}
  Since we have excluded all other possibilities, we can conclude that the remaining possibility, that the set of algorithms do not work well on Ontology data.

 \newpage
 \section{Conclusions}
 \subsection{Recap research questions}
 \subsection{Recap hypotheses}
 \subsection{Conclusion per hypothesis}
 \subsection{Overview of results}
 \subsection{Final conclusion}

 \newpage
 \section{Discussion}
 \subsection{Interpretation}
 \subsection{Unexpected results}
 \subsection{Expected results}
 \subsection{Future research}
 \subsection{Discarded implementations} %TODO consider adding to future

  \subsubsection{Just context nodes} % In future?

@@

@@

 \newpage
 \begin{appendices}
  \pagenumbering{Roman}
  \section{Appendix 1}
  \section{Appendix 2}
 \end{appendices}

>>>>

 \section*{Summary}
 %TODO somewhere else: explain why one large ontology won't work
 %TODO Definitions
 %TODO Word index
 %TODO List of tables
 %TODO List of figures
 \tableofcontents
 \newpage

%TODO \section{Motivation}

 \pagenumbering{arabic}
 \section{Introduction}
 It is a Friday afternoon and you realise you want to watch a movie with some friends this weekend. So, what do you do? To achieve this goal you must take four steps.

@@

 The most amazing thing about this story, is that it is already possible with the current technology. The algorithms to convert for example cinema website text to a computer-readable format exist. Many open databases already contain a lot of structured data that can be used to reason about data. And algorithms to combine these pieces of information to make decisions already exist and work.

@@

 \subsection{Motivation} %TODO aanleiding?
 In \cite{mccallumweblecture}, McCallum talks about Universal schema. This system learns embeddings of entities and relations found in text and Knowledge Bases. These embeddings are then used to predict new relations for the knowledge bases. The system find embeddings such that the embeddings of entities and relations that co-occur are highly similar. If some entities and relations have similar embeddings but are not known to co-occur, they are flagged as being similar. Similarly, if two entities or two relations have similar embeddings, they are likely referring to the same thing. This can be used to infer even more similarities by collaborative filtering. The technique had very promising results in Knowledge Base completion tasks.

 The embeddings are optimized by a model called the skip-gram word embedding model. It is a neural network model that learns to predict the context of a word (in this case the context of an entity is the relations it is co-occurs with and vice versa). McCallum talked about a new extension to the algorithm that allowed for homonyms and as such can deal with more ambiguous data\cite{mccallumweblecture}.

 If these embeddings can be used to predict new inter-knowledge base relations, it should also be possible to predict new intra-knowledge base relations. If these relations can be predicted, it is possible to use multiple knowledge bases together. Predicting these relations is called alignment.

 Currently, knowledge base and ontology alignment techniques do not make use of embeddings. Since embedding models have good performance on intra-knowledge base relation prediction, they should also work on ontology alignment.

 \subsection{Context}
 In this section, I will give a short summary of the context for the problem. This will enable the reader to put the problem in perspective as well as understand the reasoning behind the solution to the problem. Everything that is mentioned in this section will be explained more in-depth further on.

@@

 \subsection{Justification} %relocate?
 On the Semantic Web, many relatively small knowledge bases will be combined temporarily so that it is possible to reason over the combined knowledge of those knowledge bases. To combine these knowledge bases, their ontologies first need to be aligned. This is an act that will happen very often on the Semantic Web, since every interaction with the Semantic Web will usually require information from multiple domains. Large domains may be stored over multiple devices, which can be viewed as different knowledge bases which also need to be aligned. Therefore, it is save to say that efficient ontology matching will be very valuable.

 Since most of the knowledge on the Semantic Web is automatically gathered, it may be unreliable and even wrong. Therefore, ontology matching that is capable of handling this uncertain knowledge is required. Word Embeddings and especially multi-sense Word Embeddings should be able to handle a high amount of uncertainty.

 Also, ontology matching is already used outside of the semantic web. It is currently in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful immediately in these situations, not just in the future, on the Semantic Web.

 In summary, ontology aligning is required for multi-domain communication but poses a number of problems that make it difficult for concepts, the building blocks of ontologies, to be matched.

 \subsection{Overview} %TODO rewrite
 This thesis is structured as follows: firstly, I will describe the context of the problem, which includes the Semantic Web, Ontologies, Ontology matching and other relevant research. Secondly, I will outline the problem and research questions. Thirdly, I will describe the method I used to develop the solution, which includes the idea, challenges and considered implementations. The fourth section after this describes methods I used to evaluate the solution, with the fifth section showing the results, the sixth section the conclusions and the last and seventh section the discussion.

 \subsection{Research questions}
 The main question that I will answer with this work is the following research question:

 \subsection{Hypothesis}
 \begin{center}
 \textit{This specific matcher that uses a distributed representation of concepts can find relationships between concepts in ontologies better than the best existing matchers.}
 \end{center}

 The first three points will be shown in Section \ref{proofs}, whereas the last point is evaluated in Section \ref{conclusions}.

 This hypothesis are true for distributed representations matcher if they are true for one distributed representation matcher. Therefore to test the hypotheses, I must test them for every distinct implementation of the matcher.

\newpage
\section{Theory}
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on Ontology matching, will be discussed. Secondly, the concept of the Ontology itself will be addressed. Thirdly, current alignment strategies and their their problems are discussed. Lastly, other related research and technologies will be discussed. Later sections will refer back to the concepts discussed in this section.
 \subsection{The Semantic Web}
  The first concept is the Semantic Web, also called Web 3.0, a vision of a Web that enhances the User Experience in ways that the current Web can not provide. 
 \subsubsection{The current Web}
Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and limited interconnectivity.\footnote{For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest.} It is possible to link to content from other hubs using URLs, but this moves the user to another hub, rather than connecting the content present in the hubs.
The Semantic Web, on the other hand, contains sources of \textit{information} that can be combined whenever the user needs it. To stick with the social media example, a person can create a piece of content, say a picture of a tree, and share it with his friends. One of his friends then can share this piece of content with a group of people that likes nature pictures. However, everyone can access all references someone makes. So if a person from the nature group notices the tree is ill and comments on this, the original poster can see this comment and act accordingly. On the current web the friend who shared the picture with the group has to be contacted directly by the commenter and then has to relay the message manually.

@@

 \subsubsection{Uses of Ontologies}
 There are three main reasons Ontologies are used. Firstly, it allows querying languages to optimize the search process. Secondly, it allows expansion of information on a node. Thirdly, it allows for consistency checking a Knowledge Base. Lastly, it allows merging of knowledge bases.

@@

 \subsection{Current alignment strategies} \label{strategies}
 Many different alignment strategies have already been developed. All strategies follow the same two-step approach. The two steps are independent and as such different methods can be used interchangeably. The first step is to generate an initial correspondence set, where correspondences between nodes are found based on just their labels and meta-data. The second step is to use this initial set and the structure of the ontology to find more correspondences. These steps are called the terminological and the structural steps. Some strategies also use extra steps like the extensional and semantic steps, which use vector space models and inference, respectively\cite{future}.

@@

 Factorie is written in scala, a relatively new language that interfaces with java. See \cite{scala}. This allows for interoperability between factorie and AgreementMaker. Any algorithm that is implemented in factorie or on top of factorie can therefore be used as a matching algorithm in AgreementMaker.

@@

 \newpage
 \section{Method} %Controleer overgang
 %hier iets opschrijven?

@@

 \newpage
 \subsection{Evaluation}
 %https://hal.inria.fr/hal-00917910/document
 The goal of the research is to find an ontology alignment algorithm $A$ and prove that it performs better than current techniques. However, as current techniques may combine many different algorithms and sources of information, maybe such a comparison is not fair, nor is a comparison with individual algorithms as they may be optimised for being combined. So, instead we may want to see if performance is improved when adding $A$ to an ensemble of algorithms $E=(A_1,..,A_n)$.
 \subsubsection{Evaluation methods} \label{Evaluation methods}
 The evaluation should be done using a performance measure $P(Al,T)$ that takes the alignment an algorithm produces given an ontology and the ground truth. If both the alignment and truth are considered sets of alignments, i.e. $Al=\{a_1,..,a_n\}$ and $T=\{a'_1,..,a'_m\}$, then the True Positives $TP=\{a|a\in Al, a\in T\}$, False Positives $FP=\{a|a\in Al, a\notin T\}$ and False Negatives $FN=\{a|a\in T, a\notin Al\}$ determine the performance of the alignment. The True Negatives do not matter, as there are extremely many, and the goal of ontology matching is to find True Positives. If we only care about True Positives, the Jaccard similarity is a prime candidate for the performance measure.

@@

 It is in principle possible to weigh False Positives and False Negatives differently, but this is not done in practice. To make the results of this study easily comparable to others, equal weights will be used.
 \subsubsection{Data}
 The data used comes from a list of data sets that the OAEI provides. % Zoek paper, misschien http://www.sciencedirect.com/science/article/pii/S1570826813000188
 %Hoeveel preprocessing moest ik doen
 Every data set contains two ontologies that need to be aligned and a ground truth alignment. The goal is straightforward: find the alignment given only the ontologies and whichever outside resources are needed.

@@

 \subsection{Precision versus recall}
  \subsubsection{Mice and men data}
  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/untrained_ROC}
   \caption{The precision and recall trade-off for the results of the untrained Node2vec. $AP = 0.04194$}
   \label{ROC_untrained}
  \end{figure}

  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/Generic_ROC}
   \caption{The precision and recall trade-off for the results of the Word2vec algorithm. $AP: 0.04636$}
   \label{ROC_generic}
  \end{figure}

@@

  \begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{plots/Alternative_ROC}
   \caption{The precision and recall trade-off for the results of the alternative DeepWalk algorithm. $AP = 0.04193$}
   \label{ROC_alternative}
  \end{figure}
   \subsubsection{Medical data}
   TBD

  \subsection{F-score}
  \begin{tabular}{l|l}
  \textbf{Matcher} & \textbf{$F_1$-score}\\
  \hline
  Alternative & 0.4244\\
  Bidirectional & 0.4245\\
  Untrained & 0.4245\\
  Generic & 0.4245\\
  \textbf{AML} & \textbf{0.94}\\ %http://disi.unitn.it/~pavel/om2015/papers/om2015_proceedings.pdf#page=125
  \end{tabular}
  \subsection{Jaccard}
  TBD
  \subsection{Examples to illustrate the differences}
  TBD

 \newpage
 \subsection{Interpretation}
 \subsubsection{Unexpected results}
 \subsubsection{Expected results}

 \newpage
 \section{Conclusion and discussion}
 %\subsection{Recap research questions} % Maak 'figure' van
 %\subsection{Recap hypotheses}
 \subsection{Conclusions}
 From the results and interpretation we can conclude that distributed word representations can currently \textbf{not} find relationships between concepts in ontologies better than existing matchers.

 \subsection{Discussion}
  The system underperforms on the provided data. This can have one or more of the following four causes:
  \begin{itemize}
  \item The data breaks the considered algorithms.
  \item There are bugs in the developed system unrelated to the algorithms.
  \item The considered context extraction does not work.
  \item The considered algorithms are not suited for Ontology matching.
  \end{itemize}

  \subsubsection{The data}
  If the data set is the cause of the underperformance, logically switching to another data set will improve performance over the baseline. Possible reasons the data set causes underperformance include a too small data set or a data set having a structure that does not work with the algorithm.

  The algorithm has been run on multiple data sets with no significant difference in performance, therefore the data is not the cause of the underperformance.
  \subsubsection{The system}
  If any of the steps in the software system contains an error, the whole system may fail. This could cause all algorithms to learn badly or the results to be misinterpreted. The system consists of a data loader, preprocessor, context extractor, the algorithm, the matcher and multiple visualizers.

@@

  As such, it is safe to conclude that the system has been thoroughly tested and is not faulty.
  \subsubsection{Context extraction}
  The context extraction for DeepWalk works, as described in Section \ref{sentences}.

  The context extraction of the pre-trained Word2vec consists of splitting up every label into multiple words which can be looked up. For example "Adrenal vein" can be split up into "Adrenal" and "vein".
  The pre-trained model does not use any other form of context integration.

  For the Node2vec algorithms, the context extraction is more advanced. All edges are extracted from the Ontology using Jena. Then, for every node the context is decided as every node it has an edge with. In the case of the bidirectional algorithm, both outgoing and ingoing edges count as being connected, whereas in the monodirectional algorithm only outgoing edges count as the connectedness. For the algorithms that also use the edge labels for context, double the context size is expected.

  The number of edges is known as it has been provided by the designers of the Ontologies\cite{mice_ontology}, so we can compare the number of edges found with the actual number of edges. If these match, the context extraction works.
  \begin{table}
  \begin{tabular}{l|ll}
  & \multicolumn{2}{c}{\#edges}\\
  Algorithm & expected & actual\\
  \hline
  \textbf{Truth}\cite{mice_ontology} & \multicolumn{2}{c}{\textbf{1000000}} \\
  Monodirectional & & \\
  Bidirectional & & \\
  Bidirectional \& edges & & \\  
  \end{tabular}
  \end{table}
  Since the expected number of edges and actual number of edges match, we can conclude that the context extraction works.

  \subsubsection{The algorithm}
  Since we have excluded all other possibilities, we can conclude that the remaining possibility, that the set of algorithms do not work well on Ontology data.
 \subsection{Future research}
 \subsection{Discarded implementations}

  \subsubsection{Just context nodes}

@@

 \section{List of figures and tables}

 \newpage
 \subsection{Figures}
 \subsection{Tables}

 \section{Bibliography} %Fix

@@

 \newpage
 \section{Appendices}
 \begin{appendices}
  \pagenumbering{Roman}
  \section{Appendix 1}
  \section{Appendix 2}
 \end{appendices}
